---
title: "CH sample size for egg grant"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)

# Research Design ---------------------------------------------------------
# Two-arm parallel design
# Measurements at baseline and after intervention
# Compare change from baseline b/w the control and treatment

# Functions ---------------------------------------------------------------
# Function to calc SD of difference assuming equal variance
sd_diff <- function(sd, rho) sqrt(2 * (sd ^ 2) - 2 * rho * (sd ^ 2))

# Function to calc n based on two-sample t-test
ss_calc <- function(delta, sd) power.t.test(delta = delta, sd = sd, power = power)$n

# Assumptions -------------------------------------------------------------
# Normality of the outcome
# Equal variance b/w baseline and post-intervention
# Equal variance b/w the control and treatment
# Two-tailed with alpha = 0.05

# Parameters --------------------------------------------------------------
# Assumes 80% power
power  <- 0.80

```

## Research design

* Two-arm parallel design between a control diet and an intervention diet for 12 weeks
* Two outcomes: diastolic blood pressure (DBP) and HDL cholesterol
  * Will be measured at the baseline and at week 12
  * Compare changes from baseline between the two groups

## Assumptions

* Two-tailed tests with $\alpha = 0.05$
* For the sake of simplicity, we assumed:
  * Outcome measures are normally distributed
  * Equal variances between the baseline and the post-intervention
  * Equal variances between the control and the intervention groups

## Common parameters

* Type I error of $\alpha = 0.05$ and power of 80%
* A correlation between the baseline and the post-intervention (at week 12) was assumed to be $\rho$ = 0.4 or 0.5

## Sample size for DBP

* For DBP, we assumed:
  * A difference in mean change of 0.6, 0.8, or 1.0 mmHg between the two groups
  * A common standard deviation of 1.0 or 1.2 mmHg
  
* For a difference in mean change of 0.8 mmHg with SD = 1.2 and $\rho$ = 0.5, $n = 37$ subjects will be needed per group in order to achieve 80% of power with $\alpha = 0.05$.

```{r, echo=FALSE}
# SS calculation for BP ---------------------------------------------------
# Effect size and SD
delta <- c(0.6, 0.8, 1.0)
sd    <- c(1.0, 1.2)
rho   <- c(0.4, 0.5)

out <- expand.grid(delta = delta, sd = sd, corr = rho) %>% 
  arrange(delta, corr, sd) %>% 
  mutate(sd_change = sd_diff(sd, corr),
         n = unlist(map2(delta, sd_change, ss_calc)))
out %>% 
  mutate(sd_change = round(sd_change, 2), n = ceiling(n)) %>% 
  knitr::kable()
```

## Sample size for HDL

* For HDL, we assumed:
  * A difference in mean change of 0.025, 0.075, or 0.1 mmol/L between the two groups
  * A common standard deviation of 0.05 or 0.2 mmol/L

* For a difference in mean change of 0.025 mmol/L with SD = 0.05 and $\rho$ = 0.5, $n = 64$ subjects will be needed per group in order to achieve 80% of power with $\alpha = 0.05$.

* For a difference in mean change of 0.1 mmol/L with SD = 0.2 and $\rho$ = 0.5, $n = 64$ subjects will be needed per group in order to achieve 80% of power with $\alpha = 0.05$.

```{r, echo=FALSE}
# SS calculation for HDL --------------------------------------------------
# Effect size and SD
delta <- c(0.025, 0.075, 0.1)
sd    <- c(0.05, 0.20)
rho   <- c(0.4, 0.5)

out <- expand.grid(delta = delta, sd = sd, corr = rho) %>% 
  arrange(delta, corr, sd) %>% 
  mutate(sd_change = sd_diff(sd, corr),
         n = unlist(map2(delta, sd_change, ss_calc)))
out %>% 
  mutate(sd_change = round(sd_change, 2), n = ceiling(n)) %>% 
  knitr::kable()
```