---
title: "JO Cervical cancer additional analysis"
output: github_document
---

```{r setup, include=FALSE}

# Set working directory
setwd("M:\\Groups\\Grants_and_Studies\\AHS Cancer Screening Study")
# setwd("C:\\Users\\keiji\\Dropbox\\Consults\\Jisoo")

# Required packages
pacs <- c("tableone", "stargazer", "ggplot2", "gridExtra", "tidyverse")
sapply(pacs, require, character.only=TRUE)

# N obs = 95,873
# Data compiled by Andrew
temp <- read.csv("screening-data-file-2017-11-01.csv", header=TRUE)

###############################################
# Inclusion/exclusion criteria

# Female only
temp <- temp[temp$sex == 1 & !is.na(temp$sex),]

# Age b/w 30 and 69
temp <- temp[temp$age >= 30 & temp$age < 70 & !is.na(temp$age),]

# Own cancer, write-ins
temp$cancera <- tolower(trimws(temp$cancera))
temp$cancerb <- tolower(trimws(temp$cancerb))
temp$cancerc <- tolower(trimws(temp$cancerc))

# Search for cervical cancer
unique(agrep("cervi", temp$cancera, value=TRUE))
unique(agrep("cervi", temp$cancerb, value=TRUE))
unique(agrep("cervi", temp$cancerc, value=TRUE))

temp$prevCC <- ifelse(agrepl("cervi", temp$cancera) | 
                      agrepl("cervi", temp$cancerb) | 
                      agrepl("cervi", temp$cancerc), 1, 0)

# Exclude prevalent cases of cervical cancer
temp <- temp[temp$prevCC == 0,]

# Exclude those who had histerectomy
historec2 <- factor(temp$historec, labels=c("No", "Not sure", "Yes"))
temp <- temp[temp$historec == 1 & !is.na(temp$historec),]

# Take care of zero on test pap
temp$testpap[temp$testpap == 0] <- 1

# Exclude missing testpap (n = 793)
temp <- temp[!is.na(temp$testpap),]

# Exclude pap smear 3-4 years
# Comment or Uncomment depending on whether to include those screened 3-4 years
temp <- temp[temp$testpap != 3,]

# BMI b/w 16 and 60
ccss <- temp[temp$bmi >= 16 & temp$bmi <= 60 & !is.na(temp$bmi),]

###############################################
# Demographics and lifestyle variables

# Age as categorical
# ccss$agecat <- cut(ccss$age, c(30, 41, 51, 61, 71), right=FALSE)
ccss$agecat <- cut(ccss$age, c(30, 40, 50, 60, 70), right=FALSE)

# Race/Ethnicity
ccss$ethyou <- trimws(ccss$ethyou)
blk <- c('02', '03', '04', '05', '39', '40', '63', '74', '77', '96')
ccss$race3 <- 3
ccss$race3[ccss$ethyou == "01"]  <- 1
ccss$race3[ccss$ethyou %in% blk] <- 2
ccss$race3[ccss$ethyou %in% c("", "**")] <- NA
ccss$race3 <- factor(ccss$race3, labels=c("Non-Hisp White", "Black", "Other"))

# Marital status
ccss$marital3 <- cut(ccss$marital, c(1, 2, 5, 8), right=FALSE)
ccss$marital3 <- factor(ccss$marital3, labels=c("Never married", "Married", "Divorced/Widowed"))
ccss$marital3 <- factor(ccss$marital3, levels(ccss$marital3)[c(2, 3, 1)])

# Education
ccss$educ5 <- cut(ccss$educyou, c(1, 3, 5, 7, 8, 10), right=FALSE)
educ_levels <- c("Some HS or less", "HS/TS Diploma", "Some college", "Bachelors deg", "Postgrad deg")
ccss$educ5 <- factor(ccss$educ5, labels=educ_levels)

# Personal income
ccss$incomey[ccss$incomey >= 5] <- 5
income_levels <- c("<$10K", "$11K-20K", "$21K-30K", "$31K-50K", ">$50K")
ccss$incomey <- factor(ccss$incomey, labels=income_levels)

# Household income
ccss <- ccss %>% mutate(incomeh2 = incomeh,
                incomeh2 = factor(incomeh2, labels = c("<$10k", "$11-20k", "$21-30k", "$31-50k", "$51-75k", "$76-100k", "$101-200k", ">$200k")))
ccss$incomeh[ccss$incomeh >= 5] <- 5
ccss$incomeh <- factor(ccss$incomeh, labels=income_levels)

# BMI category
ccss$bmicat  <- cut(ccss$bmi, c(16, 18.5, 25, 30, 61), right=FALSE)
ccss$bmicat  <- factor(ccss$bmicat, labels=c("Underweight", "Normal", "Overweight", "Obese"))
ccss$bmicat2 <- factor(ccss$bmicat, levels=levels(ccss$bmicat)[c(2, 1, 3, 4)])

# Smoking
ccss$smokeEver[ccss$smoke == 1]    <- 0
ccss$smokeEver[ccss$smoke > 1]     <- 1
ccss$smokeEver[ccss$smokenow == 1] <- 1
ccss$smokeEver <- factor(ccss$smokeEver, labels=c("Never", "Ever"))

# Alcohol
ccss$alcEver[ccss$alcohol == 1] <- 0
ccss$alcEver[ccss$alcohol == 2] <- 1
ccss$alcEver[ccss$alcnow  == 1] <- 1
ccss$alcEver <- factor(ccss$alcEver, labels=c("Never", "Ever"))

# Dietary pattern
ccss$fishothf[is.na(ccss$fishothf)] <- 1
ccss$evapmlkf[is.na(ccss$evapmlkf)] <- 1 
ccss$ensuref[is.na(ccss$ensuref)]   <- 1 
ccss$ovaltinf[is.na(ccss$ovaltinf)] <- 1 

fpd <- c(0, 0.067, 0.143, 0.429, 0.786, 1, 2.5, 4.5, 6.5)
meat_var  <- c("beefhamf", "beefprof", "lambf",    "poultryf", "poulprof")
fish_var  <- c("fishsalf", "fishwhtf", "fishtunf", "fishothf")
dairy_var <- c("eggbetrf", "cheamerf", "butterf",  "mlkwholf", "mlklowff", "cottagef", 
               "checremf", "chelowff", "evapmlkf", "yogurtlf", "yogurtrf", "othdairf",
               "icecremf", "icemilkf", "ensuref",  "ovaltinf")

meat  <- Reduce("+", lapply(ccss[meat_var],  function(x) fpd[match(x, 1:9)]))
fish  <- Reduce("+", lapply(ccss[fish_var],  function(x) fpd[match(x, 1:9)]))
dairy <- Reduce("+", lapply(ccss[dairy_var], function(x) fpd[match(x, 1:9)]))
meatfish <- meat + fish

vegan <- ifelse(meat < .033 & fish < .033 & dairy < .033, 1, 0)
lacto <- ifelse(meat < .033 & fish < .033 & dairy >= .033, 1, 0)
semi  <- ifelse(meat >= .033 & (meatfish >= .033 & meatfish <= .143), 1, 0)
pesco <- ifelse(meat < .033 & fish >= .033, 1, 0)
noveg <- ifelse(meat >= .033 & meatfish > .143, 1, 0)

ccss$vegstat <- 5 * vegan + 4 * lacto + 3 * semi + 2 * pesco + 1 * noveg
ccss$vegstat <- factor(ccss$vegstat, labels=c("Non-veg", "Pesco", "Semi", "Lacto-ovo", "Vegan"))

# Physical activity
times <- ccss$walkrunf - 1
duration <- c(0, 5, 15, 25, 35, 45, 55, 60)[match(ccss$exeramt, 1:8)]
minutes  <- c(5, 15, 25, 35, 45, 55, 60)[match(ccss$walkruna, 1:7)]
distance <- c(0.25, 0.5, 1, 1.5, 2, 3, 4)[match(ccss$walkrund, 1:7)]
vigofreq <- ifelse(ccss$exerfreq - 2 < 0, 0, ccss$exerfreq - 2) 

ex4b4c <- times * distance
ex4b4d <- times * minutes
ex3a3b <- vigofreq * duration

none <- ifelse((ex3a3b == 0 & (ccss$walkrun == 2 | is.na(ccss$walkrun))) | 
               (ex3a3b == 0 & times == 0) |
               (vigofreq == 0 & ccss$walkrun == 2), 1, 0)
low  <- ifelse((0 < ex3a3b & ex3a3b < 105) | 
               (0 < ex4b4d & ex4b4d < 105) | 
               (0 < ex4b4c & ex4b4c < 3), 1, 0)
mid  <- ifelse(ex3a3b >= 105 | 
               (105 <= ex4b4d & ex4b4d < 175) | 
               (3 <= ex4b4c & ex4b4c < 9), 1, 0)
hi   <- ifelse(ex4b4d >= 175 | ex4b4c >= 9, 1, 0)

ccss$exercise[none == 1] <- 1
ccss$exercise[low  == 1] <- 2
ccss$exercise[mid  == 1] <- 3
ccss$exercise[hi   == 1] <- 4
ccss$exercise <- factor(ccss$exercise, labels=c("None", "Low", "Moderate", "Vigorous"))

###############################################
# Family cancer variables

fm_fem_ca_var <- grep("breast|ovary", names(ccss), value=TRUE)
fm_oth_ca_var <- grep("colon|prost|pancr", names(ccss), value=TRUE)
fm_all_ca_var <- c(fm_fem_ca_var, fm_oth_ca_var)
fm_ca_write   <- grep("cancero", names(ccss), value=TRUE)

# Find family history of breast or ovary cancer
fm_ca  <- Reduce("+", lapply(ccss[fm_fem_ca_var], function(x) !is.na(x)))
ccss$fmhxfemca <- ifelse(fm_ca > 0, 1, 0)

# Find among write-ins
search_words <- c('cervix', 'cervex', 'cervic', 'cevic', 'cervx', 'cervi\\.', 'crvcl', 'embryonic',
                  'endmet', 'fallop', 'fem', 'indomet', 'lungutrus', 'vagin', 'vag\\.', 'vulv', 'uter', 'womb')
search_words <- paste(search_words, collapse="|")

for (i in seq_along(ccss[fm_ca_write])){
  ccss$fmhxfemca[grep(search_words, ccss[, fm_ca_write[i]], ignore.case=TRUE)] <- 1
}

# Take care of missing
# Must remove all white spaces from write-ins
ccss[fm_ca_write] <- lapply(ccss[fm_ca_write], trimws)

# Count # of missing from write-ins
count_writein_miss <- Reduce("+", lapply(ccss[fm_ca_write], function(x) x == ""))

# COunt # of missing from all family cancers
count_circle_miss <- Reduce("+", lapply(ccss[fm_all_ca_var], is.na))

# if all missing, then assign missing
ccss$fmhxfemca[is.na(ccss$a02q5) & count_writein_miss == 10 & count_circle_miss == 25] <- NA

# Factor the variable
ccss$fmhxfemca <- factor(ccss$fmhxfemca, labels=c("No", "Yes"))

###############################################
# Pap smear
# Screening variable
ccss$screened <- ifelse(ccss$testpap == 4, 1, 0)
#ccss$screened <- ifelse(ccss$testpap %in% c(3, 4), 1, 0)

###############################################
# CCSS2 is a complete data
covars <- c("screened", "agecat", "race3", "marital3", "educ5", "incomey", "vegstat", "smokeEver", "alcEver", "exercise", "fmhxfemca")
ccss2 <- ccss[covars]
ccss2 <- ccss[complete.cases(ccss2),]
ccss2$screened2 <- factor(ccss2$screened, labels=c("Never or 5+ yrs ago", "0-2 yrs ago"))
```

## Descriptive table by dietary pattern

```{r echo = FALSE}
# Table 2 by dietary pattern
tabVars2 <- c("screened", "agecat", "age", "race3", "marital3", "educ5", "incomey", "bmicat", "bmi", 
              "smokeEver", "alcEver", "exercise", "fmhxfemca")
ccss2 %>% 
  mutate(screened = factor(screened, labels = c("Not screened", "Screened")),
         vegstat = fct_relevel(vegstat, "Pesco", after = 2)) %>% 
  CreateTableOne(tabVars2, strata="vegstat", data=.) %>% 
  print(showAllLevels = TRUE, contDigits = 1, pDigits = 4) %>% 
  kableone()

```

## Household income

* Household income (`incomeh`) has `r sum(is.na(ccss2$incomeh))` more missing values than personal income (`incomey`). We don't have a variable that indicate the number of family members in the household (`HouseHld`), though. 

```{r echo = FALSE}
# Income
# Personal vs household
ccss2 %>% select(incomey, incomeh) %>% summary()
```

* Household income in the original 8 categories:

```{r echo = FALSE}
ccss2 %>% 
  CreateTableOne("incomeh2", data=.) %>% 
  kableone()
```

## Multiple imputation by chained equations

* To evaluate a potential impact of excluding subjects with any missing covariate values, we have conducted multiple imputation (MI) as sensitivity analysis.
  * MI was conducted on the dataset prior to the exclusion of such subjects (n = 28,747), assuming covariate values are missing at random.
  * Chained equations were used to generate 5 imputed datasets using "mice" package (version 3.14.0) in R. For binary variables, logistic regression was used for missing imputation. For nominal variable with more than 2 levels, multinomial logistic regression was used. For ordinal variables, ordinal logistic regression was used. 
  * After multiple imputation, estimated beta coefficients and their standard errors from imputed datasets were pooled to obtain the final estimates using Rubin’s rule, and then prevalence ratios (PRs) were computed.

* PRs were virtually identical to those in Table 3, confirming that the exclusion of subjects with missing covariates did not affect our results.

```{r echo = FALSE}
###############################################
# Multiple imputation
library(mice)
library(geepack)

ccss_imp <- ccss[c(covars, "analysisid", "bmicat2")] %>% as_tibble()

# Imputation method
# For binal variables, use the logistic model
# For ordinal variables, use the proportional odds model ("polr")
# For nominal variables of > 2 levels, use the multinomial (polytomous) logistic model
test <- mice(ccss_imp, m = 1, maxit = 0)
method <- test$method
method[c("educ5", "incomey", "exercise")] <- "polr"

# Remove analysisid from imputation model
pred <- test$predictorMatrix
pred[, "analysisid"] <- 0

# Run multiple imputations with m = 5
ccss_mice <- mice(ccss_imp, m = 5, seed = 2948, method = method, predictorMatrix = pred)

fit <- with(ccss_mice, geeglm(screened ~ agecat + race3 + marital3 + educ5 + incomey + bmicat2 +
                                smokeEver + alcEver + vegstat + exercise + fmhxfemca, 
                              family = poisson(link = "log"), 
                              id = analysisid, 
                              corstr  = "exchangeable"))

summary(pool(fit), conf.int = TRUE) %>% 
  select(1:2, 7:8) %>% 
  mutate(across(2:4, exp)) %>% 
  mutate(across(2:4, round, 2)) %>% 
  slice(-1) %>% 
  knitr::kable()
```
